% !TEX root = ../main.tex

\chapter*{Acknowledgments}

The journey through my PhD has been one of immense learning and personal growth, facilitated by the support and guidance of many significant people. I would like to take this opportunity to express my sincere gratitude to those who have been instrumental in the completion of this dissertation and my personal journey.
 \vspace{0.5cm} 

First and foremost, I extend my deepest thanks to my supervisor, Jeremy Clark, whose expertise, patience, and insightful feedback have been invaluable throughout this research process. His guidance has been critical in shaping my work and my development as a researcher and a human. 
 \vspace{0.5cm} 
% hasty attempt to finish -> Minimum Viable PhD -> almost a dare if you know me -> 2+ months more work -> now

I am also grateful to my committee members, Ketra Schmidt, Mohammad Mannan, Amr Youssef, and Natalia Stakhanova, for their valuable suggestions that have contributed to the depth and quality of this dissertation. % TODO: update the names with the chair + ... 
 \vspace{0.5cm} 

My professional experiences have profoundly influenced my research horizon. At ConsenSys Diligence, I gained deep knowledge into Ethereum security and the composability of the decentralized ecosystem, enriching my understanding and expanding my scope. Working as the CTO at Ether Capital provided practical insights into the financial blockchain sector, further informing my research. 
 \vspace{0.5cm} 

I'm grateful to the Fe\"ytopians and my stay at the lovely Ch\^ateau du Fe\"y, where I was provided a magical place to finish this dissertation. Thank you for all the love and support. \textless3 
 \vspace{0.5cm} 
% I would also like to acknowledge the essential tools such as by Github, VSCode, \LaTeX, and ChatGPT. These resources have been crucial in facilitating my research and writing process.

Lastly, I owe a tremendous debt of gratitude to my family — my parents, my sister and her family, Hamed and Maya. Their unwavering support, encouragement, and belief in my capabilities have been my source of strength and motivation throughout this journey.
 \vspace{0.5cm} 

This dissertation stands as a testament to the collective effort and support of everyone mentioned, and for that, I am incredibly thankful.

% Be Kind,

% Shayan

